---
title: "iptw_triangulation"
author: "Darren S Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: yes
output: github_document
---
# setup

```{r setup, include=FALSE}
# configure Rmd chunks
knitr::opts_chunk$set(
    echo = TRUE,      # output code chunks
    message = FALSE,  # toggle off message output 
    warning = FALSE)  # toggle off warning output

# load frequently used packages
library(tidymodels)
library(tidyverse)
library(patchwork)

# set default ggplot theme
courier_bw <- theme_bw() +
  theme(text = element_text(family = "Courier"),
        legend.position = "bottom")

theme_set(courier_bw)

# configure connection to mysql
source("src/kale_mysql.R")
```

# outcome_regression

```{r}
# source iptw code
source("fnc/inverse_probability_treatment.R")

# source code to pre-process, run, and post-process weighted glm (for ipwt)
source("fnc/va_weighted_glm.R")
```

To run below code we need three tables for three arms:

* avastin trial arm (n 65)
* eylea ehr arm (n 4,471)
* avastin ehr arm (n 128)

```{r}
# avastin trial arm
avastin.trial <- read_csv("data/abc_patient_details.csv",
                col_types = cols(
                  id = col_character(),
                  age_at_baseline = col_integer(),
                  baseline_etdrs = col_integer(),
                  injection_count = col_integer(),
                  drug_recency = col_integer(),
                  study_exit_va = col_integer()
                )) %>% 
  rename(
    age = age_at_baseline,
    drug_load = injection_count) %>% 
  # reorder
  select(
    id,
    treatment, 
    age, 
    gender,
    baseline_etdrs,
    drug_load,
    drug_recency,
    study_exit_va) %>% 
  # restrict to avastin arm
  filter(treatment == "avastin")
```

```{sql connection = kale, output.var = "eylea.ehr"}
-- eylea ehr arm
SELECT patient_eye AS id,
treatment,
age_at_baseline AS age,
gender,
baseline_va AS baseline_etdrs,
drug_load,
drug_recency,
study_exit_va
FROM amd_synthetic_eylea_arm_study_table
WHERE eligibility = 1;
```

```{sql connection = kale, output.var = "avastin.ehr"}
-- avastin ehr arm
SELECT patient_eye AS id,
treatment,
age_at_baseline AS age,
gender,
baseline_va AS baseline_etdrs,
drug_load,
drug_recency,
study_exit_va
FROM amd_synthetic_avastin_arm_study_table
WHERE eligibility = 1;
```

```{r}
# covnert sql imports to tibbles
eylea.ehr <- as.tibble(eylea.ehr)
avastin.ehr <- as.tibble(avastin.ehr)
```


```{r}
# run ipt algorithm
iptw <- inverse_probability_treatment(
  trial_arm = avastin.ehr,   # ignore argument name
  synthetic_arm = eylea.ehr)
```

```{r}
# run primary analysis function
iptw <- iptw %>% 
  va_weighted_glm() 
```

```{r}
# ≥ 15 letters
iptw[[6]][[1]]
```

```{r}
# ≥ 10 letters
iptw[[6]][[2]]
```

```{r}
# < 15 letters
iptw[[6]][[3]]
```

Now, we calculate the concordance in the estimands between our two target trials.

```{r}
concordance <- tribble(
  ~ trial_synthetic, ~ synthetic_synthetic,
  1.55, 0.56,
  1.58, 0.67,
  1.2, 1.02
)

lm(
  formula = trial_synthetic ~ synthetic_synthetic,
  data = concordance) %>% 
  broom::glance()
```

```{r}
concordance %>% 
  ggplot(aes(x = synthetic_synthetic, y = trial_synthetic)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# delinieating_avastin_protocol

We know that Avastin was prescribed PRN during every 6 weeks the ABC trial, but how frequently during routine care?

128 of 2,056 eyes recieving 1.25 mg Avastin were eligible.

```{sql connection = kale, output.var = "avastin.injections"}
SELECT i.patient_eye, 
       i.EncounterDate AS encounter_date,
       i.InjectedDrugDesc AS injected_drug_desc,
       cumulative_injection_count,
       days_since_last_injection
FROM nvAMD_injections i
INNER JOIN amd_synthetic_avastin_arm_study_table s
ON i.patient_eye = s.patient_eye
WHERE s.eligibility = 1 AND
      i.EncounterDate > s.third_injection_date AND -- maintenance injections only
      i.EncounterDate < s.study_exit
; 
```

```{r}
avastin.injections <- avastin.injections %>% 
  as_tibble() %>% 
  arrange(
    patient_eye,
    encounter_date
  ) %>% 
  mutate(encounter_date = lubridate::ymd(encounter_date))
```

```{r}
# injection intervals vs year
avastin.injections %>% 
  mutate(encounter_year = lubridate::year(encounter_date)) %>% 
  group_by(encounter_year) %>% 
  summarise(
    mean_intervals = round(mean(days_since_last_injection), 0),
    median_intervals = round(median(days_since_last_injection), 0)
  ) %>% 
  ggplot(aes(x = as.factor(encounter_year), y = median_intervals)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 8 * 7) +
  scale_y_continuous(breaks = seq(0, 110, 10)) +
  labs(
    title = "Avastin treatment frequency in routine care is highly variable",
    x = "Calendar year of administration",
    y = "Median days since previous injection")
```