---
title: "matching_synthetic_arm_to_abc_trial_v2"
author: "Darren S Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: yes
output: github_document
---

Like `matching_synthetic_arm_to_abc_trial.Rmd`, but the propensity scoring is derived from a logistic regression fitted to abc data (i.e. probability of sbeing randomised to recieve Avastin vs. SOC) and then this model is used to calculate the ps of potential synthetic controls.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,      # output code chunks
    message = FALSE,  # toggle off message output 
    warning = FALSE)  # toggle off warning output

library(texPreview)
tex_opts$set(
  density = 600  # High resolution LaTeX output
)

# include HTML images when knitting with github_document
if (isTRUE(getOption('knitr.in.progress'))) {
  tex_opts$set(
    returnType = "html"
  )
} 
```

Load packages.

```{r message=FALSE}
library(tidymodels)
library(tidyverse)
```

Assign default ggplot theme.

```{r}
courier_bw <- theme_bw() +
  theme(text = element_text(family = "Courier"),
        legend.position = "bottom")

theme_set(courier_bw)
```

Read characteristics of abc subjects.

```{r}
abc <- read_csv("data/abc_patient_characteristics.csv",
                    col_types = cols(
                      id = col_character())) %>% 
  mutate(cohort = "abc")
```

Assign a dummy variable as being 1 if randomised to recieve Avastin and 0 if Eylea recieved.

```{r}
abc <- abc %>% 
  # create avastin_dummy that is 1 if avastin recieved and 0 if eylea
  mutate(avastin = if_else(treatment == "avastin", 1, 0)) %>% 
  rename(age = age_at_baseline,
         etdrs = baseline_etdrs)
```

# logit propensity scoring

Baseline characteristics were balanced between the ABC and EHR arms on proensity scores derived from a logistic generalised linear model regressed on treatment recieved (1 if randomised to recieve Avastin/ 0 if Eylea recieved as per standard healthcare). Baseline characteristics contributing to the pronensity scores were `baseline_va`, `age_at_baseline`, `gender` (*AND POTENTIALLY MATCHED FOR OCT THICKNESS AND PED, IF AVAILABLE*).

```{r}
logit <- glm(formula = avastin ~
               etdrs +
               age + 
               gender,
               family = binomial(link = "logit"),
             data = abc)
```

```{r echo=FALSE, results="ASIS"}
equatiomatic::extract_eq(logit,
                         use_coefs = TRUE,
                         wrap = TRUE,
                         terms_per_line = 3,
                         operator_location = "start") %>% 
  equatiomatic::preview()
```

```{r}
  augment(logit,
        data = abc,
        type.predict = "response") %>% 
  mutate(avastin = factor(avastin)) %>% 
  ggplot(aes(x = .fitted, fill = avastin)) +
  geom_density(alpha = 0.75, size = 0) +
  labs(subtitle = "Distribution of propensity scores",
       x = "Propensity score",
       y = "Density") +
  scale_fill_discrete(name = NULL, labels = c("Avastin arm", "SOC arm")) +
  scale_x_continuous(breaks = seq(0, 1, 0.1),
                     limits = c(0, 1))
```

All ABC avastin arm subejcts and potential synthetical controls were assigned a propensy score using our model trained on all ABC trial subjects. The ps distributions among the Avastin arm and potential synthetic controls was then plotted.

Read the baseline characteristics of potential synthetic controls whom recieved Eylea & combine these into a single dataframe with the baseline characteristics of those in the Avastin arm.

```{r}
ehr <- read_csv("data/ehr_patient_characteristics.csv") %>% 
  mutate(cohort = "ehr") %>% 
  rename(age = age_at_baseline,
         etdrs = baseline_etdrs)

abc.ehr <- abc %>% 
  filter(treatment == "avastin") %>% 
  select(- avastin) %>% 
  bind_rows(ehr)
```

Using our ps model to assign ps on our new dataframe.

```{r}
ps <- augment(logit,
        data = abc,
        newdata = abc.ehr,
        type.predict = "response") 

ps %>% 
  ggplot(aes(x = .fitted, fill = cohort)) +
  geom_density(alpha = 0.75, size = 0) +
  labs(subtitle = "Distribution of propensity scores",
       x = "Propensity score",
       y = "Density") +
  scale_fill_discrete(name = NULL, labels = c("ABC Avastin arm", "Potential synthetic controls")) +
  scale_x_continuous(breaks = seq(0, 1, 0.1),
                     limits = c(0, 1))
```

# singleplex psm

The *fundamental problem of causal inference* is that we know only the factual and not counterfactual outcomes.

Tenets of causal inference:

* independence---that the treatment mechanism is independent of outcome (this we assume for RCTs)
* strong ignorability (assumptions we must make for observation data) of I) unconfoundness (that treatment assignment and potential outcomes are independent of covariate attributes); and II) overlap ().


'Notice that NNM faces the risk of bad matches if the closest neighbor is far away. This issue can be resolved by imposing a tolerance level on the maximum distance,known as the caliper (see e.g., [M. Lunt. Selecting an appropriate caliper can be essential for achieving good
balance with propensity score matching. American journal of epidemiology,
179(2):226–235, 2014.]). There are some rules of thumb for choosing
the calipers (see e.g., [M. Lunt. Selecting an appropriate caliper can be essential for achieving good
balance with propensity score matching. American journal of epidemiology,
179(2):226–235, 2014.s]).

Nearest neighbour matching OR within a specified caliper without replacement using the 'tidyverse'/SQL [https://arxiv.org/pdf/1609.03540.pdf].


Try first with a caliper equal to 0.1 SD of ABC avastin propensity score.[see simulation study: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3873103/#__ffn_sectitle]
```{r}
sd(ps$.fitted[ps$treatment == 'avastin'])
```

```{r}
ps <- ps %>% 
  rename(propensity_score = .fitted) %>% 
  # add treatment dummmy variable
  mutate(treatment = if_else(treatment == "avastin", 1, 0)) %>% 
  # drop se.fit
  select(-.se.fit) %>% 
  mutate(match = "1")
```

```{r}
# randomly shuffle data before submitting to function
ps <- ps %>% 
    sample_n(size = nrow(ps), # n must be equal to n_rows
             replace = FALSE)

# propensity score matching
psm <- Matching::Match(Tr = ps$treatment, # requires verctors
                       X = ps$propensity_score, # requires verctors
                       M = 1, # 1:1 matching
                       caliper = 0.1, # units are as SD
                       replace = FALSE,
                       ties = FALSE)

# retrieve output
matches <- ps[unlist(psm[c("index.treated", "index.control")]), ]
```

```{r}
matches %>% 
  ggplot(aes(x = propensity_score)) +
  facet_grid(.~ cohort) +
  geom_histogram(alpha = 0.75, size = 0, fill = "#FF6666") +
  labs(subtitle = "Distribution of propensity scores",
       x = "Propensity score",
       y = "Density") +
  scale_x_continuous(breaks = seq(0, 1, 0.2),
                     limits = c(0, 1))
```
run other diagnostics (violin plot, box plot, AUC etc.)

How many iterations can you run before AUC tails off? Plot iterations x versus AUC on y.

# iterative matching with map
```{r}
# create a list of 10 identical dataframes
iteration.list <- list()

for (i in 1:10){ # state n iterations
  iteration.list[[i]] = ps %>% 
    # shuffle each iteration
    sample_n(size = nrow(ps), # n must be equal to n_rows
             replace = FALSE)  
  iteration.list[[i]][[10]] <-  i
}

iteration.list <- do.call(rbind.data.frame, iteration.list)

iteration.list.col <- iteration.list %>% 
  rename(iteration = V10) %>% 
  group_by(iteration) %>% 
  nest()
```

```{r}
iteration.list.col <- iteration.list.col %>% 

  mutate(psm_list = map(.x = data, ~ Matching::Match(Tr = ps$treatment,
                                                     X = ps$propensity_score,
                                                     M = 1, 
                                                     caliper = 0.1,
                                                     replace = FALSE,
                                                     ties = FALSE))) %>% 
  # error in extracting matches
  mutate(matches = map(.x = psm_list, 
                       ~ unlist(psm[c("index.treated", "index.control")]),))
```

gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))))

Issues:

* returning same controls for each iteration. You may need to create your own for loop/function that changes seed per iteration.
* mutate(matches = map(.x = psm_list, ~ unlist(psm[c("index.treated", "index.control")]),))for [[4]] is a list of ids and not dataframe with associated with ids ()

# iteration with bespoke function

```{r}
ps <- ps %>% 
  mutate(match_counter = 0)
```

shuffle between each iteration (necessary to set new seed?)
exclude those already matched from subsequent iterations

```{r}
# shuffled as as row in list column & 50 (set new seed per iteration)

```

Play around with the algortihms to find that which best balances covariates/ps (use `Matching::MatchBalance`)

* check the balance of ps, and each of the coavriates included in the ps model


AFTER MATCHING, AUC BETWEEN ABC & EHR PS DISTRIBUTIONS SHOULD BE ~ 0.5. INDEED, COAVRIATES INCLDUED IN THE PROPENSITY SCORE SHOIULD BE MATCHED.

CALCULATE 'VARIANCE OF THE EFFECT ESTIATES'.

# Analysis (compare two Loess curveS?)

adjust for injection_count in the analysis.