---
title: "exchangeability_plot"
author: "Darren S Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: yes
output: github_document
---

```{r setup, include=FALSE}
# configure Rmd chunks
knitr::opts_chunk$set(
    echo = TRUE,      # output code chunks
    message = FALSE,  # toggle off message output 
    warning = FALSE)  # toggle off warning output

# load frequently used packages
library(broom)
library(tidyverse)

# set default ggplot theme
courier_bw <- theme_bw() +
  theme(text = element_text(family = "Courier"),
        legend.position = "bottom",
        strip.background = element_rect(
          colour = "black", 
          fill = "white",
          size = 0.5))

theme_set(courier_bw)
```

```{r}
# create list of filenames to read from
y <- list(
  "../data/cohort_nc.csv",
  "../data/cohort_iptw.csv",
  "../data/cohort_em.csv",
  "../data/cohort_psm.csv"
)

# read cohorts from .csv
z <- map(
  .x = y,
  ~ read_csv(
    file = .x,
    col_types = cols(
      treatment = col_factor(levels = c('eylea', 'avastin'))
    ))
)

# name each element of list
names(z) <- c('nc', 'iptw', 'em', 'psm')
```

```{r}
# merge to a single tbl
zz <- bind_rows(z, .id = "cohort") %>% 
  # 
  select(cohort:baseline_etdrs)
```

```{r}
# create assign_ps fnc
assign_ps <- function(
  tbl
){

  # source propensity model
  source("../src/propensity_model.R")

  # predict Pr of treatment(Avastin == 1)
  output <- augment(
   x = propensity_model,
   newdata = tbl,
   type.predict = "response") %>% 
    rename(
      ps = .fitted,
      ps_se = .se.fit
    )
  
  # calculate ipw for iptw cohhort, else weight = 1
  output <- output %>% 
    mutate(weight = if_else(cohort == "iptw", 1 / ps, 1))
  
  output
  }
  
```

```{r}
# score Pr
exc <- assign_ps(tbl = zz)
```

```{r}
exc %>% 
  ggplot(aes(x = ps, colour = treatment, weight = weight)) +
  facet_wrap(~ cohort) +
  geom_density()
```

# add auc

```{r}
# calculate auc for each method
auc <- exc %>% 
  group_by(cohort) %>% 
  yardstick::roc_auc(
    ps,
    truth = treatment
  )
```

```{r}
# calculate weighted auc

w.roc <- WeightedROC::WeightedROC(
    guess = exc$ps[exc$cohort == "iptw"], 
    label = exc$treatment[exc$cohort == "iptw"],
    weight = exc$weight[exc$cohort == "iptw"])

WeightedROC::WeightedAUC(w.roc)

```

```{r}
# replace unweighted auc with weighted auc

w.auc <- tribble(
  ~ cohort, ~ .estimate,
  "iptw", WeightedROC::WeightedAUC(w.roc)
)

auc <- auc %>% 
  filter(cohort != "iptw") %>% 
  bind_rows(w.auc)

```

```{r}
exc %>% 
  mutate(cohort = factor(
    cohort, 
    levels = c("nc", "iptw", "em", "psm"))) %>% 
  ggplot(aes(x = ps)) +
  facet_grid(
    cohort ~.,
    labeller = labeller(cohort = c(
      "nc" = "NC",
      "iptw" = "IPTW", 
      "em" = "EM", 
      "psm" = "PSM"))) +
  geom_density(aes(colour = treatment, weight = weight)) +
  geom_text(
    data = auc, 
    aes(
      x = 0.8, 
      y = 4, 
      label = paste("AUC", format(round(.estimate, 3), nsmall = 3))),
    family = "Courier") +
  labs(
    x = "Propensity score",
    y = "Density"
  )
```

'Inferences about treatment effect made using propensity-score matching are valid only if, in the matched sample, treated and untreated subjects have similar distributions of measured baseline covariates.'

# legend to generic names, remove legend title.
# facet order acting up
# second graph is standardised mean differences in age, sex (or categorical equiavlent), baseline_read see (PMC3472075), or add them to table 1. https://handbook-5-1.cochrane.org/chapter_9/9_2_3_2_the_standardized_mean_difference.html

```{r}
exc %>% 
  ggplot(aes(x = treatment, y = ps, weight = weight, colour = treatment)) +
  facet_wrap(~ cohort) +
  geom_violin() +
  geom_boxplot(width = 0.15)
  

```

