---
title: "exact_matching_eylea"
author: "Darren S Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: yes
output: github_document
---
# Setup

Script for exact matching synthetic controls to ABC Avastin arm eyes (n 65).

Output tables export to kale are:

* matched.va as [eylea_exact_va_long]
* matched.pairs.long as [eylea_exact_pairs_long]

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,      # output code chunks
    message = FALSE,  # toggle off message output 
    warning = FALSE)  # toggle off warning output
```

Load frequently used packages.

```{r}

```

Setup mysql connection.

```{r}
kale <- DBI::dbConnect(RMySQL::MySQL(),
                        user = "admin", 
                        password = "password",
                        dbname = "NOVA6",
                        host = "127.0.0.1",
                        port = 9999)
```

Set new default ggplot theme.

```{r}
courier_bw <- theme_bw() +
  theme(text = element_text(family = "Courier"),
        legend.position = "bottom")

theme_set(courier_bw)
```

Read abc data.

```{sql connection = kale, output.var = "abc"}
SELECT id, 
       treatment,
       gender,
       injection_count, 
       age_at_baseline, 
       baseline_etdrs
FROM abc_patient_details;
```

```{r}
abc <- abc %>% 
  mutate(gender_m = if_else(gender == "M", 1, 0)) %>% 
  filter(treatment == "avastin") %>% 
  rename(
    age = age_at_baseline) %>% 
  unite(col = "age_gender_etdrs", 
        c("age", "gender", "baseline_etdrs"), 
        sep = "-",
        remove = FALSE) %>% 
  select(id,
         age_gender_etdrs, 
         gender_m,
         age, 
         baseline_etdrs)
```

Import pool of potential synthtic controls from kale.

```{sql connection = kale, output.var = "ehr"}
SELECT patient_eye AS id,
       injection_count,
       gender,
       age_at_baseline AS age,
       baseline_va AS baseline_etdrs
FROM amd_synthetic_eylea_arm_study_table   
WHERE eligibility = 1;
```

Convert gender to a dummy variable.

```{r}
ehr <- ehr %>% 
  mutate(
    gender_m = if_else(gender == "M", 1, 0)) %>% 
  select(id,
         gender_m,
         age,
         baseline_etdrs)
```

Convert to list-column workflow

```{r}
abc.nest <- abc %>% 
  group_by(id, age_gender_etdrs) %>% 
  nest()
```

Create a new list column with all exact matched.

```{r}
abc.nest <- abc.nest %>% 
  mutate(matches = map(.x = data, 
                       ~ left_join(.x,
                                   ehr,
                                   by = c("gender_m",
                                          "age",
                                          "baseline_etdrs"),
                                   keep = TRUE)))
```

Randomly sample 1 control for each ABC subject.

```{r}
set.seed(1337)

abc.nest <- abc.nest %>% 
  mutate(sampled_control = map(.x = matches,
                               ~ sample_n(.x, size = 1)))
```

Unnest list column.

```{r}
abc.unnest <- abc.nest %>% 
  select(sampled_control) %>% 
  unnest() %>% 
  ungroup() %>% 
  rename(
    abc_id = id,
    ehr_id = id1)
```

Create look-up table for matched pairs.

```{r}
matched.pairs <- abc.unnest %>% 
  drop_na(ehr_id) %>% 
  select(age_gender_etdrs,
         abc_id,
         ehr_id) %>% 
  # add pair_id
  mutate(pair_id = 1:43) # change 43 to nrows()

# create the same table in long form
matched.pairs.long <- matched.pairs %>% 
  pivot_longer(cols = abc_id: ehr_id,
               names_to = "cohort",
               values_to = "id")
```

Export matched IDs (`matched.pairs.long`) to Kale.

```{r}
DBI::dbWriteTable(
  conn = kale,
  name = "eylea_exact_pairs_long",
  value = matched.pairs.long,
  overwrite = TRUE
)
```

# Exploration of unmatched eyes.

Forty-three of 65 (0.66) eyes had exact matches. The 21 subjects that had no exact matches were:

```{r}
abc.unnest %>%
  filter(is.na(ehr_id)) %>% 
  select(abc_id,
         age_gender_etdrs) %>% 
  arrange(age_gender_etdrs)
```

Create a table with 43 matched ABC subjects & 43 matched controls + va measurements in long form for plotting.

```{r}
matched.abc <- abc %>% 
  filter(id %in% matched.pairs$abc_id) %>% 
  mutate(cohort = "abc") %>% 
  # add pair_id
  left_join(select(matched.pairs, abc_id, pair_id),
            by = c("id" = "abc_id"))

matched.ehr <- abc.unnest %>% 
  drop_na(ehr_id) %>% 
  mutate(cohort = "ehr") %>% 
  rename(id = ehr_id) %>% 
  select(- abc_id) %>% 
  # add pair_id
  left_join(select(matched.pairs, ehr_id, pair_id),
            by = c("id" = "ehr_id"))

# merge tables
matched <- bind_rows(matched.abc, matched.ehr)
```

Explore those who had no exact matches.

```{r}
abc.unmatched <- abc %>% 
  anti_join(matched.abc,
            by = "id")
abc.unmatched
```

```{r}
abc.unmatched %>% 
  pivot_longer(cols = gender_m: baseline_etdrs,
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(aes(x = value, fill = variable)) +
    facet_wrap(~ variable,
               scales = "free_x") +
    geom_histogram() +
    scale_y_continuous(breaks = seq(0, 12, 2))
```

Low n but distributions of age and baseline_etdrs appear not to be overly skewed---though summary statstics and distibution plots are unlikely to be telling given the limited sample n.

```{r}
abc.unmatched %>% 
  summarise(mean_age = floor(mean(age)),
            sd_age = round(sd(age), 1),
            mean_baseline_etdrs = round(mean(baseline_etdrs), 0),
            sd_baseline_etdrs = round(sd(baseline_etdrs), 1))
```

# LOESS BELOW.

Read abc va data in longform.

```{sql connection = kale, output.var = "abc.va"}
SELECT id, week, etdrs
FROM abc_va_longform
WHERE id IN (
  SELECT id 
  FROM eylea_exact_pairs_long
);

```

```{r}
abc.va <- abc.va %>% 
  mutate(cohort = "abc")
```

Import all va measurements during study period for all potential synthetic controls.

```{sql connection = kale, output.var = "ehr.va"}
SELECT s.patient_eye AS id,
       v.EncounterDate,
       s.baseline_eylea_date,
       DATEDIFF(v.EncounterDate, s.baseline_eylea_date) AS days_since_baseline,
       CEIL(DATEDIFF(v.EncounterDate, s.baseline_eylea_date) / 7) AS week,
       v.max_etdrs AS etdrs
FROM amd_synthetic_eylea_arm_study_table s
LEFT JOIN nvAMD_visual_acuity v
ON s.patient_eye = v.patient_eye
WHERE s.eligibility = 1 AND
     v.EncounterDate = s.baseline_va_date OR
     v.EncounterDate > s.baseline_eylea_date AND
     v.EncounterDate <= s.study_exit;
     
-- verbose, but the clunky WHERE clauses pulls all baseline va measurements taken before date of first injection 
```

Whittle down tibble for va measures for just those exact matched.

```{r}
ehr.va <- ehr.va %>% 
  as.tibble() %>% 
  filter(id %in% matched.pairs$ehr_id) %>% 
  mutate(cohort = "ehr")
```

Bind together `abc.va` & `ehr.va`.

```{r}
matched.va <- abc.va %>% 
  rbind(select(ehr.va, id, week, etdrs, cohort))

# Impute ETDRS <0 as 0
matched.va$etdrs[matched.va$etdrs < 0] <- 0
# convert cohort to drug recieved
matched.va$drug[matched.va$cohort == "abc"] <- "Avastin (1.25 mg q6w)"
matched.va$drug[matched.va$cohort == "ehr"] <- "Eylea (2 mg q8w)"
```

Export `matched.va` to kale.

```{r}
DBI::dbWriteTable(
  conn = kale,
  name = "eylea_exact_va_long",
  value = matched.va,
  overwrite = TRUE
)
```

SCRIPT END.